---
title: "Grip strength, mating success, and immune function in US adults: a registered report "
author: "C. B. Smith and E. H. Hagen"
date: 03/23/2023
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(ggplot2)
library(dplyr)
library(psych)
library(xtable)
library(knitr)
library(kableExtra)
library(weights)
library(jtools)
library(reshape2)
library(gtsummary)
library(corrplot)
source('draft_analysis.R')
```
EHB Research Report

##Abstract

Evidence suggests that the mating benefits of muscle mass in human males trade off with costs of increased energy intake and decreased measures of native immunity, likely due to an evolutionary history of sexual selection. It is unknown if females experience a similar tradeoff. Using data from the 2011-2012 phase of the National Health and Nutrition Examination Survey, a large representative sample of the US, we will test whether grip strength, a proxy for upper body strength, is a positive predictor of self-reported outcome measures of adult mating success (lifetime number of sexual partners, past year number of sexual partners, and age at first intercourse) in both males and females, and negatively related to a measure of native immune function, white blood cell count, controlling for numerous anthropometric, socioeconomic, hormone, health, and physical activity related confounds. These results replicate an earlier study of US males. However, there was either no significant sex difference in the effect of grip strength, or the effect in females was smaller than males but still positive, evidence that a similar trade off in the costs and benefits of muscle exists for women.



```{r summarytable}
d_G[d_G$age>=18 & d_G$age<=60,]  %>% select("sex", "age", "race", "strength", "bmi", "partnered", "edu", "height", "weight", "whitebloodcell", "testosterone") %>%
  tbl_summary(by = sex,
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} / {N} ({p}%)"),
              digits = all_continuous() ~ 2,
        #      type = list(where(is.numeric) ~ "continuous"),
             # type = list(partnered ~ "categorical"),
              type = all_dichotomous() ~ "categorical",
              label  = list(
                age = "Age (Years)",
                race = "Race and Ethnicity",
                strength = "Combined Grip Strength",
                bmi = "BMI (kg/m^2)",
                partnered = "Partnered Status",
                edu = "Education", 
                height = "Height (cm)", 
                weight = "Weight (kg)", 
                whitebloodcell = "WBCC (1000 cells/µL)", 
                testosterone = "Testosterone (ng/dL)",
                age_first_sex = "Age at first sex (Years)", 
                sex_partners = "Lifetime number of sexual partners", 
                sex_partners_year = "Past year number of sex partners"), 
              missing_text = "Missing"
  ) %>% add_p() %>%
  add_overall() %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Sex**") %>% 
    modify_caption("**Table 1. Participant Characteristics**")



# designsG$d.design.adults %>%
#   tbl_svysummary(by = sex, include = c(age, edu, maritalstatus, income, race, bmi, 
#                              whitebloodcell, strength, testosterone, partnered, 
#                              height, weight, special_equipment)) 
```

```{r cor, fig.width=12, fig.height=12 }
# cordat <- d_G[, c("age", "sex1", "strength", "bmi")]
# cordat <- cordat[cordat$age>=18 & cordat$age<60,]
# cormat <- round(cor(cordat, use = "complete.obs"),2)



cor_mat <- svycor(
  ~ sex_partners +
    sex_partners_year +
    age_first_sex +
    age +
    strength +
    bmi +
    edu +
    hemoglobin +
    whitebloodcell +
    depression +
    chronic_disease_score +
    physical_disease_count +
    testosterone +
    vigorous_rec +
    moderate_rec +
    vigorous_work +
    moderate_work,
  designsG$d.design.adults,
  na.rm = T
)

cor_mat_cors <- cor_mat$cors
corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75)

# meltedcor_mat <- melt(cor_mat_cors)
# 
# ggplot(data = meltedcor_mat, aes(x=Var1, y=Var2, fill = value)) +
#   geom_tile()


```




```{r exact, fig.width = 12, fig.height = 12}

m_lifetime <- svyglm(
  sex_partners ~
    age_centered * sex +
    strength_centered * sex +
    partnered  +
    bmi_centered * sex,
  family = quasipoisson(),
  design = designsG$d.design.adults
)

m_pastyear <- svyglm(
  sex_partners_year ~
      age_centered * sex +
      strength_centered * sex +
      partnered * strength_centered +
     bmi_centered * sex,
    family = quasipoisson(),
    design = designsG$d.design.adults
  )

m_agefirst <-  svyglm(
    age_first_sex ~
      age_centered * sex +
      strength_centered * sex +
      partnered  +
      bmi_centered * sex,
    family = gaussian(),
    design = designsG$d.design.adults
  )


mnames1 <- c(
  "Age at first intercourse",
  "Lifetime Number of Sexual Partners",
  "Past Year Number of Sexual Partners"
)

vnames1 <- c(
  "strength_centered" = "Strength (S)",
  "sexfemale" = "Sex (Female)",
  "sexfemale:strength_centered" = "Sex (Female) x Strength",
  "age_centered" = "Age (S)",
  "partneredTRUE" = "Partnered",
  "strength_centered:partneredTRUE" = "Partnered x Strength (S)",
  "age_centered:sexfemale" = "Age x Sex (Female)",
  "height_centered" = "Height (S)",
  "weight_centered" = "Weight (S)",
  "bmi_centered" = "BMI (S)",
  "sexfemale:bmi_centered" = "Sex (Female) x BMI"
)

forestplot(m_agefirst, m_lifetime, m_pastyear, intercept = F, facet = F, dodgewidth = .5, modelnames = mnames1,
  varnames = vnames1)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .5)) +
labs(title = "Exact Models: Regression coefficients from generalized linear models", color = "Model") + guides(colour = guide_legend(reverse = T))

#visualize coefs from models derived from lassek and gaulin. 
#add coef table below to match l and g 
#decrease dodgewidth, reorder models to match points 

```




```{r table2_exact}

mm1 <- tbl_regression(m_lifetime, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)


mm2 <- tbl_regression(m_pastyear, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI"
                      )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

mm3 <- tbl_regression(m_agefirst, 
                      label = list(
                        age_centered = "Age",
                        sex = "Sex",
                        strength_centered = "Grip strength",
                        partnered = "Currently partnered",
                        bmi_centered = "BMI", 
                        edu = "Education",
                        total_work_MET = "Occupational exertion"
    #              label = list(
    # age_centered = "Age",
    # sex = "Sex",
    # strength_centered = "Grip strength",
    # partnered = " Currently partnered", 
    # bmi_centered = "BMI"
    # edu = "Education",
    # total_work_MET = "Occupational exertion"
  )) %>%
  add_significance_stars(
    hide_se = TRUE,
    pattern = "{estimate}{stars}<br>({std.error})"
  ) %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

  tbl_merge(
    tbls = list(mm1, mm2, mm3),
     tab_spanner = c("Lifetime number of sexual partners",
                     "Past year number of sexual partners",
                     "Age at first sexual intercourse")) %>% 
  as_gt() %>%
  gt::tab_header(title = "Table 2. Exact Models of Mating Success",
                 subtitle = "Standardized regression coefficents from generalized
                 linear models for three indices of mating success for men and women ages
                 18-59. Models derived from Lassek and Gaulin 2009")

```

```{r lifetime, fig.width = 12, fig.height = 42}

#include groups of potential confounds to explore lack of sex effect for all 4 mating outcomes 

fig1 <- forestplot(
  manth1, msoc1, mheal1, mphys1, mhor1a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames, varnames = vnames)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Lifetime Number of Sexual Partners", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig2 <- forestplot(
  manth2,msoc2,mheal2, mphys2, mhor2a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames
)$plot + theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(.8)) +
  labs(title = "Past Year Number of Sexual Partners", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig3 <- forestplot(
  manth3, msoc3, mheal3, mphys3, mhor3a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames)$plot +
  theme_minimal(20) +
  geom_pointrange(size = 1.2, position = position_dodge(width = .7)) +
  labs(title = "Age at First Sex", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig4 <- forestplot(
  manth4, msoc4, mheal4, mphys4, mhor4a,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  modelnames = mnames,
  varnames = vnames)$plot +
  theme_minimal(20) + geom_pointrange(size = 1.2, position = position_dodge(width = .8)) +
  labs(title = "Currently Partnered", color = "Controls") +
  coord_cartesian(clip = "off")
# + xlim(-2.5, 3.55)

fig1234 <- ggarrange(fig1, fig2, fig3, fig4,
                     labels = "AUTO",
                     font.label = list(size = 20),
                     ncol = 1, nrow = 4,
                     common.legend = TRUE, legend="bottom")

 annotate_figure(fig1234,
                       top = text_grob("Figure 2. Regression coefficients from generalized linear models of mating success\n",
                                       size = 20, face = "bold",
                                       hjust = 0,
                                       x = 0.01),
                       bottom = text_grob("\nVariables labelled (S) have been standarized by 2 SD. Base level of race/ethnicity is Mexican American.",
                                          size = 18))

```

```{r table2_mod}
# 
# mm1_m <- tbl_regression(mm1_mod, 
#                         label = list(
#                           age_centered = "Age",
#                           sex = "Sex",
#                           strength_centered = "Grip strength",
#                           partnered = "Partnered",
#                           median_salary_current = "Median Salary",
#                           age_first_sex = "Age at first sex"
#                         ))  %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
# mm2_m <- tbl_regression(mm2_mod, 
#                         label = list(
#                           sex = "Sex",
#                           strength_centered = "Grip strength",
#                           partnered = "Partnered",
#                           bmi_centered = "BMI"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
#     
# mm3_m <- tbl_regression(mm3_mod, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           total_work_MET = "Occupational exertion",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
#   tbl_merge(
#     tbls = list(mm1_m, mm2_m, mm3_m),
#     tab_spanner = c("Lifetime number of sexual partners",
#                     "Past year number of sexual partners",
#                     "Age at first sexual intercourse")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 3. Modified Models of Mating Success",
#                  subtitle = "Standardized regression coefficents from generalized
#                  linear models for three indicies of mating success. Models derived
#                  from exploratory data")

```


```{r table2_vaginalsex}

# mm1_v <- tbl_regression(m1_vaginal_sex_partners, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
# mm2_v <- tbl_regression(m2_vaginal_pastyear, 
#                         label = list(
#                           age_centered = "Age",
#                           edu = "Education",
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
#   
#   tbl_merge(
#     tbls = list(mm1_v, mm2_v),
#     tab_spanner = c("Lifetime number of vaginal sex partners",
#                     "Past year number of vaginal sex partners")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 4. Models of Mating Success (Vaginal Sex)",
#                  subtitle = "Standardized regression coefficents from generalized
#                  linear models for lifetime and past year vaginal sex partners")

```


```{r immunefig}
fig5 <- forestplot(
  mwbc,
  intercept = F,
  facet = F,
  dodgewidth = .8,
  varnames = vnames)$plot +
  theme_minimal(25) +
  geom_pointrange(size = 2, position = position_dodge(width = .8)) +
  labs(title = "White blood cell count") +
  # labs(title = "Figure.3 Regression coefficients from generalized linear model of immune investment.",
  #      subtitle = "White Blood Cell Count",
  #      caption = "Variables with (S) have been standarized by 2 SD.") +
  theme(legend.position = "none")

 annotate_figure(fig5,
                           top = text_grob("Figure 3. Regression coefficients from generalized linear model of immune investment\n",
                                           size = 20, face = "bold",
                                           hjust = 0,
                                           x = 0.01),
                           bottom = text_grob("Variables labelled (S) have been standarized by 2 SD.",
                                              size = 18))

```

Regression coefficients for predictors of WBC count in men and women 18-59, exact model from lassek and gaulin

```{r immunemodel}

tbl_regression(
  mwbc,
  label = list(
    age_centered = "Age",
    bmi_centered = "BMI",
    sex = "Sex",
    strength_centered = "Grip strength"
  )
) %>%
  add_significance_stars(hide_se = TRUE,
                         pattern = "{estimate}{stars}<br>({std.error})") %>%
  modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
  modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)

# m_lymph <- tbl_regression(m_l, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# m_mono <- tbl_regression(m_m, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# m_neutro <- m_wbc <- tbl_regression(m_n, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# 
# m_eos <- tbl_regression(m_e, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# m_baso <- tbl_regression(m_b, 
#                         label = list(
#                           bmi = "BMI", 
#                           sex = "Sex",
#                           strength_centered = "Grip strength"
#                         )) %>%
#   add_significance_stars(
#     hide_se = TRUE,
#     pattern = "{estimate}{stars}<br>({std.error})"
#   ) %>%
#   modify_header(label ~ "**Variable**", estimate ~ "**Beta (SE)**") %>%
#   modify_footnote(estimate ~ "SE = Standard Error", abbreviation = TRUE)
# 
# tbl_merge(
#     tbls = list(m_wbc, m_lymph, m_mono, m_neutro, m_eos, m_baso),
#     tab_spanner = c("White blood cell count",
#                     "Lymphocyte count", 
#                     "Monocyte count", 
#                     "Neutrophil count", 
#                     "Eosinophil count", 
#                     "Basophil count")) %>% 
#    as_gt() %>%
#   gt::tab_header(title = "Table 4. Models of Immune Function",
#                  subtitle = "Regression coefficents from generalized
#                  linear models of white blood cell count, and cell type counts")


```


